### 2022-06-06 README.Rmd
paprpkgr::new_paper()

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Copeland.2022.hypoxia.flux

<!-- badges: start -->
<!-- badges: end -->

### Compendium URL

### Author
William Oldham (wm.oldham@gmail.com)

### Contents
This research compendium contains all data, code, and text associated with our publication:

>

The paper is available:

### Use

### Licenses

| **Text / figures :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
| **Code :** See the [DESCRIPTION](DESCRIPTION) file.
| **Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/). Attribution requested in reuse.