### 2022-06-06 analysis/qbias-correction-factors.Rmd

---
title: "Q Bias Correction Factors"
subtitle: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  bookdown::pdf_document2:
    number_sections: true
    toc: true
    toc_depth: 1
    df_print: kable
    dev: cairo_pdf
    latex_engine: xelatex
geometry: margin = 0.5in
fontsize: 11pt
mainfont: Calibri
---

```{r setup, include = FALSE}
# load libraries
suppressPackageStartupMessages({
  devtools::load_all()
  library(tidyverse)
  library(wmo)
  library(targets)
  library(mzrtools)
})

# resolve conflicts
conflicted::conflict_prefer("filter", "dplyr")

# set global chunk options
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = TRUE, 
  fig.align = "center", 
  out.width = "49%"
)

options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 
    "latex" else "html"
})

theme_set(theme_wmo(base_family = "Calibri"))
```

```{r}
withr::with_dir(here::here(), {
  qbias_ratios <- tar_read(qbias_ratios)
  correction_factors <- tar_read(correction_factors)
})
```

\newpage

# Overview

A selected ion monitoring method was utilized to increase sensitivity for isotope detection. Others have noted quadrupole bias on the Q Exactive using this detection method (*i.e.*, the sensitivity for each isotope differs relative to its position within the m/*z* window). To account for this, naturally labeled standards were analyzed using methods where the m/*z* window is shifted -1.003355 units, the ^13^C mass defect.

# Analysis

The M1/M0 ratio was calculated for each frame shift. The resulting data were fit with a third-degree polynomial function.

```{r}
make_plots <- function(data, batch, metabolite) {
  ggplot(data) +
    aes(x = window, y = ratio) +
    geom_smooth(
      method = MASS::rlm, 
      formula = y ~ poly(x, 3),
      color = "blue", 
      se = FALSE
    ) +
    geom_point(
      size = 2, 
      alpha = 0.5, 
      color = "black"
    ) +
    geom_point(
      size = 5, 
      stat = "summary", 
      fun.data = "mean_se", 
      color = "blue"
    ) +
    labs(
      title = str_c(metabolite, batch, sep = " "), 
      x = "Window", 
      y = "Ratio"
    ) 
}
```

```{r, results = "hide", out.width = "24.9%", fig.show = "hold"}
qbias_ratios |> 
  nest() |> 
  mutate(plots = pmap(list(data, batch, metabolite), make_plots)) |> 
  pull(plots)
```

```{r}
qbias_ratios |> 
  select(-ratio) |> 
  pivot_longer(c(M0, M1), names_to = "isotope", values_to = "value") |> 
  mutate(
    M = str_extract(isotope, "\\d") |> as.numeric(), 
    M = M + window, 
    iso = str_c("M", M)
  ) |> 
  left_join(correction_factors, by = c("metabolite", "batch", "iso" = "M")) |> 
  mutate(value = value * cf) |> 
  select(-c(cf, iso, M)) |> 
  pivot_wider(names_from = isotope, values_from = value) |> 
  mutate(ratio = M1 / M0) |> 
  group_by(metabolite, batch) |> 
  nest() |> 
  mutate(plots = pmap(list(data, batch, metabolite), make_plots)) |> 
  pull(plots)
```

```{r}
usethis::use_data(correction_factors, overwrite = TRUE)
```