### 2022-08-03 README.Rmd
render README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Copeland.2022.hypoxia.flux

<!-- badges: start -->
<!-- badges: end -->

### Author
William Oldham (woldham@bwh.harvard.edu)

### Contents
This research compendium contains all data, code, and text associated with our publication:

> **MYC overrides HIF to regulate proliferating primary cell metabolism in hypoxia.** Courtney A. Copeland, Benjamin A. Olenchock, David Ziehr, Sarah McGarrity, Kevin Leahy, Jamey D. Young, Joseph Loscalzo, William M. Oldham. *bioRxiv* 2020.09.21.306464. https://doi.org/10.1101/2020.09.21.306464.

### Licenses

| **Text / figures :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
| **Code :** See the [DESCRIPTION](DESCRIPTION) file.
| **Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/). Attribution requested in reuse.